### 2016-03-29 Manuscript/BiologyLetters/Missing_mammals-draft.tex
Table latex
Updated minor reviews from reviewer 2 and 4 + title change (reviewer 3)

\noindent{\Large \bf Assessment of available anatomical characters for phylogenetic analysis among living mammals}%Assessment of cladistic data availability for living mammals

@@

%---------------------------------------------
%
%       INTRODUCTION
%
%---------------------------------------------
\newpage 
\section{Introduction}
There is an increasing consensus among biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013}.
To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
One increasingly popular method, the Total Evidence method \cite{ronquista2012}, ccombines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method (e.g. \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
A downside of this method is that it requires molecular data for living taxa and morphological/anatomical data for both living and fossil taxa.
Sections of this data can be difficult, or impossible, to collect for every taxon in the analysis.
For example, fossils rarely have molecular data and incomplete fossil preservation may reduce the number of anatomical characters available.
Additionally, it has become less common to collect anatomical characters for living taxa when molecular data is available (e.g. in \cite{slaterphylogenetic2013}, only 13\% of living taxa have coded anatomical characters).
Unfortunately this missing data can lead to errors in phylogenetic inference.
Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between coded anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
This is because (1) fossils will not be placed accurately within the correct clade if it contains no coded anatomical characters for living taxa; and (2) fossils have a higher probability of being placed within clades with more coded anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.

@@

%\input{Table_results}

@@

%\begin{figure}[!htbp]
%\centering
%    \includegraphics[width=1\textwidth]{figure1.pdf}
%\caption{Phylogenetic distribution of species with available coded anatomical characters across two orders (A: Primates; B: Carnivora).
%Blue branches indicate species with available coded anatomical characters.}
%\label{Figure_example_coverage}
%\end{figure}

@@

When few species have available coded anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil being placed in the correct clade.
The second best scenario is that species with available characters are randomly distributed across the phylogeny. 
Here we expect no bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available coded anatomical characters were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with available characters are phylogenetically clustered. 
Then we expect two major biases: first, fossils will not be placed within a clade containing no data, and second, fossils will have higher probability of being placed within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade because they have no coded anatomical characters available.
Instead the fossil will have a high probability of being placed on a branch that contains many anatomical characters such as within the Canidae or Ursidae (Figure \ref{Figure_example_coverage}B). 
This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of herpestids as having zero-length branches for anatomical characters, and canids and ursids having long branches and thus "attracting" fossil placements.

It is worth noting, however, that our analysis did not include all the matrices containing anatomical characters ever published.
In fact, our data collection procedure focused on including studies that provided matrices easily accessible, i.e. we did specifically not include any matrices that were only available in paper format (e.g. printed in books), non-reusable format (e.g. an image of the matrix)  or/and matrices available only upon request (e.g. by emailing the authors).
The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high.

Despite the absence of good morphological/anatomical data coverage for living mammals, the Total Evidence method still seems to be the most promising way of combining living and fossil species in macroevolutionary analyses.
Following the recommendations in \cite{GuillermeCooper}, we must code anatomical characters for as many living species as possible. 
Fortunately, mammal specimens are usually readily available in natural history collections, therefore, we propose increased effort into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
Such efforts would be valuable not only to phylogeneticists, but also to any researcher focusing understanding macroevolutionary patterns and processes.

>>>>

\noindent{\Large \bf Assessment of available anatomical characters for linking living mammals to fossil taxa in phylogenetic analyses}
%TG: or actually, along the same lines:
%"Assessment of available anatomical characters for linking living and fossil mammals in phylogenetic analyses"
%TG: What do you think?

@@

%---------------------------------------------
%
%       INTRODUCTION
%
%---------------------------------------------
\newpage 
\section{Introduction}
There is an increasing consensus among biologists that studying both living and fossil taxa is essential for fully understanding macroevolutionary patterns and processes \cite{slaterunifying2013,fritzdiversity2013}.
To perform such analyses it is necessary to combine living and fossil taxa in phylogenetic trees.
One increasingly popular method, the Total Evidence method \cite{ronquista2012}, combines molecular data from living taxa and morphological data from both living and fossil taxa in a supermatrix that can then be used with the tip-dating method (e.g. \cite{pyrondivergence2011,ronquista2012,schragocombining2013,slaterunifying2013,beckancient2014}), producing a chronogram with living and fossil taxa at the tips.
A downside of this method is that it requires molecular data for living taxa and discrete morphological/anatomical data for both living and fossil taxa.
Sections of this data can be difficult, or impossible, to collect for every taxon in the analysis.
For example, fossils rarely have molecular data and incomplete fossil preservation may reduce the number of anatomical characters available.
Additionally, it has become less common to collect anatomical characters for living taxa when molecular data is available (e.g. in \cite{slaterphylogenetic2013}, only 13\% of living taxa have coded anatomical characters).
Unfortunately this missing data can lead to errors in phylogenetic inference.
Simulations show that the ability of the Total Evidence method to recover the correct topology decreases when there is little overlap between coded anatomical characters in living and fossil taxa, and that the effect of missing data on topology is greatest when living taxa have few anatomical characters available \cite{GuillermeCooper}.
We might expect the total evidence method to perform poorly when there is little overlap between coded anatomical characters in living and fossil taxa.
This is because fossil taxa cannot be correctly placed within a clade of living species if none of its members have been coded for morphological characters.
Furthermore, simulations show that fossils are more likely to be placed in clades with for which more characters have been coded, regardless of whether this is the correct clade \cite{GuillermeCooper}.
% This is because (1) fossils will not be placed accurately within the correct clade if it contains no coded anatomical characters for living taxa; and (2) fossils have a higher probability of being placed within clades with more coded anatomical characters available for living taxa, regardless of whether this is the correct clade \cite{GuillermeCooper}.

@@

\input{Table_results}

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{figure1.pdf}
\caption{Phylogenetic distribution of species with available coded anatomical characters across two orders (A: Primates; B: Carnivora).
Blue branches indicate species with available coded anatomical characters.}
\label{Figure_example_coverage}
\end{figure}

@@

When few species have available coded anatomical characters, the ideal scenario is for them to be evenly distributed (as measured by phylogenetic overdispersion) to maximize the possibilities of a fossil being placed in the correct clade.
The second best scenario is that species with available characters are randomly distributed across the phylogeny. 
Here we expect no bias in the placement of fossils \cite{GuillermeCooper}, it is therefore encouraging that for most orders, species with available coded anatomical characters were randomly distributed across the phylogeny.
The worst case scenario for fossil placement is that species with available characters are phylogenetically clustered. 
Then we expect two major biases: first, fossils will not be placed within a clade containing no data, and second, fossils will have higher probability of being placed within the most sampled clade by chance. 
Our results suggest that this may be problematic at the genus-level in Carnivora, Chiroptera and Soricomorpha. 
%For example, a Carnivora fossil will be unable to be placed in the Herpestidae clade because they have no coded anatomical characters available.
For example, a carnivoran fossil is unlikely to be placed in Herpestidae because they have no coded anatomical characters available.
Instead the fossil will have a high probability of being placed on a branch that contains many anatomical characters such as within the Canidae or Ursidae (Figure \ref{Figure_example_coverage}B). 
This is analogous to the problem of long-branch attraction/short branch repulsion, as one can think of herpestids as having zero-length branches for anatomical characters, and canids and ursids having long branches and thus "attracting" fossil placements.

We acknowledge, however, that our analysis did not include all the matrices containing anatomical characters ever published.
In fact, our data collection procedure focused on including studies that provided easily accessible matrices, i.e. we did specifically not include any matrices that were only available in paper format (e.g. printed in books), non-reusable format (e.g. an image of the matrix)  or/and matrices available only upon request (e.g. by emailing the authors).
The difficulty of obtaining such matrices makes them less likely to be included in Total Evidence analyses, especially where the number of taxa to include in the phylogeny is high.

Despite the absence of good morphological/anatomical data coverage for living mammals, the Total Evidence method still seems to be the most promising way of combining living and fossil species in macroevolutionary analyses.
Following the recommendations in \cite{GuillermeCooper}, we must code anatomical characters for as many living species as possible. 
Fortunately, mammal specimens are usually readily available in natural history collections, therefore, we propose increased effort into coding anatomical characters from living species, possibly by engaging in collaborative data collection projects.
Such efforts would be valuable not only to phylogeneticists, but also to any researcher focusing on understanding macroevolutionary patterns and processes.
